---
title: "Practical Machine Learning"
author: "Brett Taylor"
date: "September 15, 2015"
output: html_document
---
What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).
2. You should also apply your machine learning algorithm to the 20 test cases available in the test data above. Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details. 


Outcome: classe

#Steps    
1. How did I build the model
2. How did I use cross validation
3. What do I think the expected out of sample error is
4. Why did I make the choices
5. Predict 20 cases

#Building the predictor model

##Choosing the predictor variables
First note that there a lot of variables that only have NA's in place or very few values so remove them because the will be very little value in predicting.

##Clean the Data
There are many variables in this data set that have a large number of missing values.   

```{r cleanData}
library(dplyr)
cleanData <- function (data) {
  obsCount <- nrow(data)
  tnames <- names(data)
  results <- c()
  index <- 1
  for(i in 1:ncol(data)){
    numberOfNAs <-sum(is.na(data[,i]))
    
    if(numberOfNAs > (.5 * obsCount)) {
      results[index] <- (tnames[i])
      index <- index + 1
    }
  }
  
  data <- (data[,setdiff(tnames,results)])
 

  return(data)
}
pmltraining <- read.csv("./data/pml-training.csv",na.strings = c("","#DIV/0!","NA"))
pmltraining <- cleanData(pmltraining)
pmltraining <-  pmltraining%>%select(-Column1)
pmltesting <-read.csv("./data/pml-testing.csv",na.strings = c("","#DIV/0!","NA"))
pmltesting <- cleanData(pmltesting)


```

##Build the Model
```{r }
library(caret)
library(dplyr)
set.seed(2040)
pmltraining <- pmltraining%>%filter(classe != 4)
 inTrain <- createDataPartition(y = pmltraining$classe,  p = .75, list = FALSE) 
 training <- pmltraining[inTrain,]
 validation <- pmltraining[-inTrain,]

## The format of the results 

```

##Create a Model      
```{r}
 fitControl <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           ## repeated twenty times
                           repeats = 3)
  plsFit <- train(classe ~ ., 
                  data = training, 
                  method = "pls",
                  trControl = fitControl,
                  preProc = c("center", "scale"))
  
    gmbFit1 <- train(classe ~ ., 
                  data = training, 
                  method = "gbm",
                  trControl = fitControl
                  )
str(gmbFit1)
 
```

##Predict 

```{r}
validation.results <- predict(gmbFit1,newdata = validation)
validation.df <- data.frame(validation$classe,validation.results)
denom <- nrow(validation.df)
matchCount  <- sum(validataion.df[,1] == validation.df[,2])
accuracyOfData <- (matchCount/denom)*100
confusionMatrix(validation.df$validation.results,validation.df$validation.classe)

```

##Testing Set

```{r}







testing.results <- predict(gmbFit1,newdata = pmltesting)

```